### 2018-04-02 MSDS6372_Final_Project_BBHOF.Rmd
Added some text in the Intro and PCA

---
title: "MSDS6372 - Final Project"
author: "Benjamin Wilke, Michael Landrum"
date: "March 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# required libraries
library(kableExtra) # to make uber sexy tables for output
library(pastecs) # for easy descriptive statistics
library(ggbiplot) # implementation of biplot using ggplot2 for plotting PCs
library(Amelia) # cool library for exploring missing data
library(ROCR) # for ROC plots and AUC calculations
# turn off scientfic notation for the entire script
options(scipen = 999)
```

## Introduction

Many players in the National Baseball League (MLB) become eligible for the National Baseball Hall of Fame after they have played 10 seasons of MLB and have been retired for five full seasons. For example, those players eligible for consideration in 2019 will have played their final game in 2013. However, many players who become eligible are not admitted.

This project will explore the career performance statistics that influence whether a player will be admitted into the National Baseball Hall of Fame. The core data set was procured from The American Statistical Association (http://ww2.amstat.org/publications/jse/datasets/MLBHOF.new.txt) and includes data assembled from The Baseball Encyclopedia (Reichler 1993) and Total Baseball (Thorn and Parmer 1993).

..

Finally, once we've established a model for determining whether a player will be admitted to the National Baseball Hall of Fame from the core data - we will make predictions on players that were not in the core data.

## Exploratory Data Analysis

```{r, echo=TRUE}
# import the data
BBHOFRaw <- read.table("MLBHOF-tab.new.dat.txt", sep="\t", header=FALSE)

# create column names
colnames(BBHOFRaw) <- c("Name", "NumSeasonsPlayed", "GamesPlayed", "OfficialAtBats", "RunsScored", "Hits", "Doubles", "Triples", "HomeRuns", "RunsBattedIn", "Walks", "Strikeouts", "BattingAvg", "OnBasePct", "SluggingPct", "AdjustedProduction", "BattingRuns", "AdjustedBattingRuns", "RunsCreated", "StolenBases", "CaughtStealing", "StolenBaseRuns", "FieldingAvg", "FieldingRuns", "PrimaryPositionPlayed", "TotalPlayerRating", "HOFMembership")

# fix some import typecasting issues and apply NA null for values with "."
BBHOFRaw$Name <- as.character(BBHOFRaw$Name)

BBHOFRaw$Strikeouts[BBHOFRaw$Strikeouts == "."] <- NA
BBHOFRaw$Strikeouts <- as.integer(BBHOFRaw$Strikeouts)

BBHOFRaw$StolenBases[BBHOFRaw$StolenBases == "."] <- NA
BBHOFRaw$StolenBases <- as.integer(BBHOFRaw$StolenBases)

BBHOFRaw$CaughtStealing[BBHOFRaw$CaughtStealing == "."] <- NA
BBHOFRaw$CaughtStealing <- as.integer(BBHOFRaw$CaughtStealing)

BBHOFRaw$StolenBaseRuns[BBHOFRaw$StolenBaseRuns == "."] <- NA
BBHOFRaw$StolenBaseRuns <- as.integer(BBHOFRaw$StolenBaseRuns)

# display top 5 observations
knitr::kable(BBHOFRaw[1:5,],caption = "Top 5 Observations from Baseball Hall of Fame Data", row.names = FALSE, "html") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"))
```

# Missing Values in the Raw Data
The following is a graphical representation of the completeness of the data. It is also followed with a table of counts of NAs in the data. Fortunately, this data is very complete and the only features that are affected by NAs are: StolenBaseRuns, CaughtStealing, Strikeouts, and StolenBases.

```{r, echo=TRUE, fig.width=8, fig.height=10}
# display missing values map (graphic)
missmap(BBHOFRaw, main="Missing Values in Raw Data")

# calculate NAs for raw data, apply row labels as new column
NAdataframe <- sapply(BBHOFRaw, function(x) sum(is.na(x)))

# display NAs in kable table
knitr::kable(NAdataframe,caption = "Counts of NAs (Missing Values) in Raw Data", row.names = TRUE, "html") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width=FALSE, position = "left")
```

#

```{r, echo=TRUE}
# display frequency table for did/did not get inducted into HOF
BBHOFRaw$HOFMembership[BBHOFRaw$HOFMembership == 2] <- 1# since the labels also include 2, converting those to 1, so the result is truly binary (1, 0)
BBHOFRaw$HOFMembership[BBHOFRaw$HOFMembership == 2] <- 1
count(BBHOFRaw, "HOFMembership")

# descriptive statistics, load into new data frame for processing
descriptiveTable <- stat.desc(BBHOFRaw)

# remove non-numeric features
dropcolumns <- c("Name","PrimaryPositionPlayed", "HOFMembership")
descriptiveTable <- descriptiveTable[,!(colnames(descriptiveTable) %in% dropcolumns)]

# round all numeric values to 2 decimal points
descriptiveTable <- round(descriptiveTable, 2)

# remove rows for statistics we don't care about, leaving: N, Mean, Median, Std Dev, Var, Min, Max
remove <- c("nbr.null", "nbr.na", "range", "sum", "SE.mean", "CI.mean", "coef.var")
descriptiveTable <- descriptiveTable[-which(rownames(descriptiveTable) %in% remove),]

# display descriptive statistics
knitr::kable(descriptiveTable,caption = "Descriptive Statistics for Numeric Features in the Baseball Hall of Fame Data", row.names = TRUE, "html") %>%
  kable_styling(bootstrap_options = c("striped","hover", "condensed", "responsive"))

```

# Principal Components Analysis

```{r echo=TRUE}

# remove non-numeric features and features with missing values (NEED TO REVISIT THIS)
dropcolumnsPCA <- c("Name", "PrimaryPositionPlayed", "HOFMembership", "Strikeouts", "StolenBases", "CaughtStealing", "StolenBaseRuns")
BBHOF_PCA <- BBHOFRaw[,!(colnames(BBHOFRaw) %in% dropcolumnsPCA)]

# extract labels (player made it into HOF), convert to factor
BBHOF_Labels_PCA <- BBHOFRaw["HOFMembership"]
BBHOF_Labels_PCA$HOFMembership <- as.factor(BBHOF_Labels_PCA$HOFMembership)

# perform PCA
BBHOF_PCA <- prcomp(BBHOF_PCA, center=TRUE, scale.=TRUE)

# print principal components (only the first 4 PCs)
print(BBHOF_PCA$rotation[,1:4])

# summary of principal components
summary(BBHOF_PCA)

```

```{r, echo=TRUE, fig.width=15, fig.height=10}
# plot PCs using ggbiplot
ggbiplot(BBHOF_PCA, obs.scale=1, var.scale=1, groups=BBHOF_Labels_PCA$HOFMembership, ellipse=TRUE, circle=TRUE) + theme(legend.position='top') + scale_color_discrete()

```

alright....working some logistic regression..

```{r, echo=TRUE}

# copy data set, remove Name from the raw data set, so it's not included in the model, also removing features to play with the model
BBHOF_logistic <- BBHOFRaw
dropcolumns <- c("Name", "StolenBaseRuns", "StrikeOuts", "CaughtStealing", "StolenBases")
BBHOF_logistic <- BBHOF_logistic[,!(colnames(BBHOF_logistic) %in% dropcolumns)]

# split the raw data into testing and training data
set.seed(100) # set seed so that same sample can be reproduced in future

# now selecting 75% of data as sample from total 'n' rows of the data  
sample <- sample.int(n=nrow(BBHOF_logistic), size=floor(.75*nrow(BBHOF_logistic)), replace=FALSE)

# subset the data using the sample integer vector created above
train <- BBHOF_logistic[sample, ]
test  <- BBHOF_logistic[-sample, ]

# fitting the binomial logistic regression model, HOFMembership is dependent, fitting using all features
model <- glm(train$HOFMembership ~., family=binomial(link='logit'),data=train)

summary(model)

# predict based on the test data, type='response' output probabilities in the form of P(y=1|X)
fittedresults <- predict(model, newdata=test, type='response')

# there are some NAs in the fitted results...we need to check into that - not sure how that's possible if I'm removing the features with NAs
sum(is.na(fittedresults))

# if P(y=1|X) > 0.5 then y = 1 otherwise y=0
fittedresults <- ifelse(fittedresults > 0.5, 1, 0)

# calculate the mean of the fitted results that don't equal the observed result - IGNORE NAs
misClasificError <- mean(fittedresults != test$HOFMembership, na.rm=TRUE) # this adds up all the instances of misclassification then divides by total (via mean)

# print the output as 100% - error
print(paste('Accuracy',1-misClasificError))

```

and now some ROC curves...

```{r, echo=TRUE}

pr <- prediction(fittedresults, test$HOFMembership)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc

```

Future Eligible: https://baseballhall.org/hall-of-famers/future-eligibles